My title

### 2024-11-29 paper/paper.qmd
EDA and model for predicting wickets

library(tidyverse)
library(arrow)
library(here)
library(tibble)
library(knitr)
cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
cleaned_data <- cleaned_data %>% filter(year >= 2021)
```

@@

# Data and Graphs

@@

stadium_boundaries <- cleaned_data %>%
  group_by(venue) %>%
  summarise(
    num_boundaries = sum(is_boundary == TRUE),  
    num_balls = n(),                             
    boundary_pct = num_boundaries / num_balls
  ) %>% arrange(desc(boundary_pct), desc(num_balls))

ggplot(stadium_boundaries, aes(x = reorder(venue, boundary_pct), y = boundary_pct)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Boundary Percentage by Stadium",
       x = "Stadium Name",
       y = "Boundary Percentage") +
  theme_minimal() +
  coord_flip()

@@

target_boundaries <- cleaned_data %>%
  group_by(match_id, target) %>%
  summarise(
    num_boundaries = sum(is_boundary == TRUE),  
    num_balls = n(),                             
    boundary_pct = num_boundaries / num_balls
  ) %>% arrange(desc(boundary_pct), desc(num_balls))

ggplot(target_boundaries, aes(x = target, y = boundary_pct)) +
  geom_point(color = "steelblue", size = 3, alpha = 0.5) +
  labs(title = "Boundary Percentage vs Target per Match",
       x = "Target",
       y = "Boundary Percentage") +
  geom_smooth(method = "lm", color = "red") +
  theme_minimal() 
```

over_boundaries <- cleaned_data %>%
  group_by(over) %>%
  summarise(
    num_boundaries = sum(is_boundary == TRUE),  
    num_balls = n(),                             
    boundary_pct = num_boundaries / num_balls
  ) %>% arrange(desc(boundary_pct), desc(num_balls))

ggplot(over_boundaries, aes(x = over, y = boundary_pct)) +
  geom_point(color = "steelblue", size = 3) +
  labs(title = "Boundary Percentage vs Over Number",
       x = "Over",
       y = "Boundary Percentage") +
  theme_minimal()

@@

bowling_batting_matchup_boundaries <- cleaned_data %>%
  group_by(bowling_style, batting_style) %>%
  summarise(
    num_boundaries = sum(is_boundary == TRUE),  
    num_balls = n(),                             
    boundary_pct = num_boundaries / num_balls
  ) %>% arrange(desc(boundary_pct), desc(num_balls))

ggplot(bowling_batting_matchup_boundaries, aes(x = bowling_style, y = boundary_pct, fill = batting_style)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Boundary Percentage by Bowling and Batting Style",
       x = "Bowling and Batting Style",
       y = "Boundary Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

@@

over_bowlers <- cleaned_data %>%
  group_by(over, bowling_style) %>%
  summarise(ball_count = n())
```

```{r}
model <- glm(is_boundary ~ target, family = binomial, data = cleaned_data)
summary(model)
```
```{r}

model <- glm(is_boundary ~ over, family = binomial, data = cleaned_data)
summary(model)

model <- glm(is_boundary ~ target + bowling_style + batting_style, family = binomial, data = cleaned_data)
summary(model)

@@

@@

>>>>

library(tidyverse)
library(arrow)
library(here)
library(tibble)
library(knitr)
library(modelsummary)
library(broom)
library(marginaleffects)

cleaned_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
test_data <- read_parquet(here("data/02-analysis_data/test_data.parquet"))
set.seed(888)
```

@@

# Data

## Measurement

## Predictor Variables

num_bowlers_per_type <- cleaned_data %>%
  group_by(bowling_style) %>% 
  summarise(
    num_bowlers = n_distinct(bowler)                            
  )

num_batters_per_type <- cleaned_data %>%
  group_by(batting_style) %>% 
  summarise(
    num_bowlers = n_distinct(striker)                          
  )

## Relationship Between Wickets and other Variables

@@

stadium_boundaries <- cleaned_data %>%
  group_by(venue) %>%
  summarise(
    num_matches = n_distinct(match_id),
    num_wickets = sum(wicket == TRUE),  
  ) %>% arrange(desc(num_wickets), desc(num_matches))

ggplot(stadium_boundaries, aes(x = venue, y = (num_wickets/num_matches))) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
       x = "Stadium Name",
       y = "Wickets Per Match") +
  theme_minimal() +
  coord_flip()

@@

over_boundaries <- cleaned_data %>%
  group_by(over) %>%
  summarise(
    num_wickets = sum(wicket == TRUE),  
    num_balls = n()
  ) %>% arrange(desc(num_wickets), desc(num_balls))

ggplot(over_boundaries, aes(x = over, y = num_wickets)) +
  geom_point(color = "steelblue", size = 3) +
  labs(
       x = "Over",
       y = "Number of Wickets") +
  theme_minimal()

@@

bowling_batting_matchup_boundaries <- cleaned_data %>%
  group_by(bowling_style, batting_style) %>%
  summarise(
    num_wickets = sum(wicket == TRUE),  
    num_balls = n(),                             
  ) %>% arrange(desc(num_wickets), desc(num_balls))

ggplot(bowling_batting_matchup_boundaries, aes(x = bowling_style, y = num_wickets, fill = batting_style)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
       x = "Bowling Style",
       y = "Number of Wickets") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}
wickets_prev_over_wickets <- cleaned_data %>%
  group_by(over) %>%
  summarise(
    num_wickets = sum(wicket == TRUE),
    prev_over_wickets = mean(prev_over_wickets),
    num_balls = n(),                             
  ) %>% arrange(desc(num_wickets), desc(num_balls))

ggplot(wickets_prev_over_wickets, aes(x = prev_over_wickets, y = num_wickets)) +
  geom_point(color = "steelblue", size = 3, alpha = 0.7) +
  labs(
       x = "Wickets in Previous Over Wickets",
       y = "Number of Wickets") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(cleaned_data, aes(x = run_rate, fill = factor(wicket))) +
  geom_histogram(position = "identity", alpha = 0.8, bins = 30) +
  labs(
    x = "Run Rate",
    y = "Count",
    title = "Histogram of Run Rate by Wicket Occurrence"
  ) +
  scale_fill_discrete(name = "Wicket Occurred", labels = c("No Wicket", "Wicket")) + 
  theme_minimal()
```

## Model set-up

### Model justification

# Results

```{r}
simple_glm_wicket_model <- readRDS(here("models/simple_glm_wicket_model.rds"))

#summary(simple_glm_wicket_model)
modelsummary(simple_glm_wicket_model)
```

```{r}
complex_glm_wicket_model <- readRDS(here("models/complex_glm_wicket_model.rds"))
#summary(complex_glm_wicket_model)
modelsummary(complex_glm_wicket_model)
```

```{r}
overly_complex_glm_wicket_model <- readRDS(here("models/overly_complex_glm_wicket_model.rds"))
#summary(overly_complex_glm_wicket_model)
modelsummary(overly_complex_glm_wicket_model)
```

# Simple Model Summary

```{r}
overly_complex_glm_wicket_model <- complex_glm_wicket_model <- readRDS(here("models/overly_complex_glm_wicket_model.rds"))

#summary(overly_complex_glm_wicket_model)
modelsummary(overly_complex_glm_wicket_model)
```

```{r}
simple_glm_wicket_model_predictions <-
  predictions(simple_glm_wicket_model) |>
  as_tibble()

simple_glm_wicket_model_predictions |>
  mutate(wicket = factor(wicket)) |>
  ggplot(aes(x = over, y = estimate, color = wicket)) +
  stat_ecdf(geom = "point", alpha = 0.75) +
  labs(
    x = "The Over",
    y = "Estimated Probability that a wicket will occur",
    color = "Was actually a wicket"
  ) +
  theme_classic() +
  theme(legend.position = "bottom")
```

```{r}
test_data_simple <- test_data
predictions <- predict(simple_glm_wicket_model, newdata = test_data_simple, type = "response")

test_data_simple$predicted_wicket_prob <- predictions
test_data_simple <- test_data_simple %>% 
  mutate(predicted_wicket = predicted_wicket_prob >= 0.5) %>% 
  mutate(correct_prediction = predicted_wicket == wicket)

summary_results <- test_data_simple %>% group_by(wicket) %>% 
  summarise(
  correct = sum(correct_prediction),
  incorrect = sum(!correct_prediction)
)

summary_results

## Complex Model Summary

```{r}
complex_glm_wicket_model_predictions <-
  predictions(complex_glm_wicket_model) |>
  as_tibble()

complex_glm_wicket_model_predictions |>
  mutate(wicket = factor(wicket)) |>
  ggplot(aes(x = prev_over_wickets, y = estimate, color = wicket)) +
  stat_ecdf(geom = "point", alpha = 0.75) +
  labs(
    x = "The Number of Wickets Taken in the Previous Over",
    y = "Estimated Probability that a wicket will occur",
    color = "Was actually a wicket"
  ) +
  theme_classic() +
  theme(legend.position = "bottom")
```

```{r}
test_data_complex <- test_data
predictions <- predict(complex_glm_wicket_model, newdata = test_data_complex, type = "response")

test_data_simple$predicted_wicket_prob <- predictions
test_data_simple <- test_data_simple %>% 
  mutate(predicted_wicket = predicted_wicket_prob >= 0.5) %>% 
  mutate(correct_prediction = predicted_wicket == wicket)

summary_results <- test_data_simple %>% group_by(wicket) %>% 
  summarise(
  correct = sum(correct_prediction),
  incorrect = sum(!correct_prediction)
)

summary_results
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\appendix

# Appendix {-}

# Additional data details

@@

bowling_batting_role_matchup_boundaries <- cleaned_data %>%
  group_by(bowling_style, batter_playing_role) %>%
  summarise(
    num_wickets = sum(wicket == TRUE),  
    num_balls = n(),                             
  ) %>% arrange(bowling_style, batter_playing_role)

@@

# Model details {#sec-model-details}

## Posterior predictive check

@@

What to cite:
- cricketdata
- ESPNCricinfo
- Cricsheet
- kable
- knitr
- R
- tidyverse
- ggplot
